\documentclass{iutbscthesis}
\usepackage{kantlipsum} % This package is used to generate place holder text. Remove it in the actual world.

\tolerance=1 % Allow some flexibility for line breaking
\emergencystretch=\maxdimen % Allow line breaking to stretch a little more
\hyphenpenalty=10000 % Allow hyphenation with a reasonable penalty
\hbadness=10000 % Suppress most warnings about underfull or overfull lines

% The iutbscthesis uses biblatex for bibliography management.
% You can learn the basics of biblatex from:
% https://www.overleaf.com/learn/latex/Bibliography_management_with_biblatex
\addbibresource{citations.bib}
\graphicspath{{./figures/}}
\DeclareGraphicsExtensions{.pdf, .png, .jpg, .jpeg}


% The complete title of your thesis, mandatory
\title{\Large{Optimizing CAM Refinement with Swin-based Feature Affinity for Weakly Supervised Semantic Segmentation}}

% Use the \addauthor command to keep adding authors with student ID
\addauthor{Anika Farzana}{200041204}
\addauthor{K. M. Abesh Ahsan}{200041225}
\addauthor{Sayemah Amin}{200041234}

% Use the \supervisor command to define the supervisor with name, designation, and department
% There needs to be one and only one supervisor
\supervisor{Dr. Md. Hasanul Kabir}{Professor}{Department of Computer Science and Engineering}

% Use the \addcosupervisor command to add a cosupervisor. There might be no cosupervisor at all
% % In that case, just remove all of the \addcosupervisor commands.
% \addcosupervisor{Dr.\ Sue Permann}{Assistant Professor}{Department of Computer Science and Engineering}
% \addcosupervisor{Mr. Francis FullOfFrenchPeople}{Lecturer}{Department of Computer Science and Engineering}

% Complete department name, mandatory
\department{Department of Computer Science and Engineering}

% Complete program name, mandatory
\program{BACHELOR OF SCIENCE IN COMPUTER SCIENCE AND ENGINEERING}

% Date (day, month name, and year) when the presentation took place, mandatory
\defensedate{29}{September}{2025}

\begin{document}

%-------------------------- FRONT MATTER --------------------------%

% Generate cover page of the report to be printed on the hard cover
\coverpage

% Switch to roman page numbering
\pagenumbering{roman}

% Generate the title leaf, mandatory
\titlepage

% Auto generate declaration of candidate based on previous info, mandatory
\declarationofcandidate

% % Dedicate report to someone. You may add more sentences to it.
\dedicatedto{We dedicate this thesis to our parents, whose constant love, guidance, and support have been the foundation of our achievements and the inspiration behind our efforts.}

\tableofcontents
\listoffigures
\listoftables

\clearpage

% Add abbreviations within this environment using the \abbr command
% \begin{abbreviations}
%     \abbr{CNN}{Convolutional Neural Network}
%     \abbr{PIP}{PIP Installs Packages}
%     \abbr{TikZ}{TikZ ist kein Zeichenprogramm}
%     \abbr{WIKI}{What I Know Is}
% \end{abbreviations}

% Any acknowledgement that you want to write goes in the following environment

\begin{acknowledgement}
    I am profoundly grateful to my supervisor, Dr. Md. Hasanul Kabir, Professor, Department of Computer Science and Engineering, for his outstanding guidance, support, and feedback throughout this research journey. His expertise and encouragement have been pivotal in shaping the direction and quality of this thesis.
    
    His patience and dedication to his work have inspired me to aim for excellence and approach challenges with curiosity and determination. His constructive feedback and insightful suggestions have significantly deepened my understanding of the domain and inspired me to think critically and creatively.
    
    It has been an honor to work under his supervision, and I deeply appreciate his invaluable time and effort in guiding me through this research.
\end{acknowledgement}

% Write your abstract here.
\begin{abstract}
    \addcontentsline{toc}{chapter}{Abstract}
    Semantic segmentation is a fundamental computer vision task requiring pixel-level understanding of images. While fully supervised methods achieve high accuracy, they rely on costly pixel-level annotations. Weakly Supervised Semantic Segmentation (WSSS) mitigates this by using weaker supervision, such as image-level labels, to train effective models. Recent WSSS progress leverages Class Activation Maps (CAMs), though their sparsity and poor boundary localization remain challenges. This study enhances CAM quality through multi-modal backbones like UniCL and hierarchical transformers such as Swin Transformer for stronger feature extraction, coupled with an affinity-based framework that fuses encoder and decoder affinities for semantically coherent pseudo-labels. A Pixel-Adaptive Refinement (PAR) module further improves object boundaries using local similarity cues. Experiments on the PASCAL VOC 2012 dataset yield mean IoUs of 50.3\% (validation) and 50.8\% (test), with strong performance on large, distinctive classes but weaker results for small or human-centric ones due to CAM bias and dataset imbalance. Overall, our findings demonstrate that UniCL and Swin Transformer significantly improve CAM quality and segmentation under weak supervision while underscoring the need for scale-aware, bias-compensated refinement strategies.
\end{abstract}


%-------------------------- MAIN BODY --------------------------%
% Switch to arabic page numbering
\pagenumbering{arabic}

\input{chapters/introduction/inroduction.tex}
\input{chapters/relatedworks/related-works.tex}
\input{chapters/methodology/methodology.tex}
\input{chapters/result_analysis/result_analysis.tex}
\input{chapters/Discussion/discussion.tex}
\input{chapters/conclusion/conclusion.tex}

% \input{chapters/citations/citations.tex}

\end{document}
